---
title: "HW 6"
author: "Olga Lyashevskaya, George Moroz, Alla Tambovtseva and Ilya Schurov"
output:
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(tidyverse)
theme_set(theme_bw())
```

There are several tasks on

* data manipulation with `dplyr` (if you have any problems see [this](https://www.datacamp.com/courses/dplyr-data-manipulation-r-tutorial) and [this](https://www.datacamp.com/courses/joining-data-in-r-with-dplyr) courses)
* data visualisation with `ggplot2` (if you have any problems see [this course](https://www.datacamp.com/courses/data-visualization-with-ggplot2-1))

### 1.
These datasets contain the results of the experiment that evaluates the inter speaker variation in noun class assignment among speakers of the Zilo dialect of Andi (a Nakh-Daghestanian language). In Zilo there are two classes for inanimate objects with no obvious semantic distinction between them.

There are two datasets:

* [zilo_class_experiment_data.csv](https://raw.githubusercontent.com/LingData2019/LingData/master/data/zilo_class_experiment_data.csv)
  * w_id --- word id
  * stimulus --- stimuli word in Zilo
  * translation_en --- translation to English
  * stimulus_source --- the source of the stimulus: native or loan
  * 1 ... 16 --- columns that contain answers of 16 speakers of Zilo: `b` or `r` (class markers)
* [zilo_class_experiment_informants.csv](https://raw.githubusercontent.com/LingData2019/LingData/master/data/zilo_class_experiment_informants.csv)
  * s_id --- speaker id (corresponds to numbers in the previous dataset)
  * sex --- sex of the speaker
  * age_2017 --- age of the speaker on the moment of the interview

### 1.1
Calculate the ratio of b-words ($\frac{b\text{-words}}{b\text{-words}+r\text{-words}}$) for each sex and stimulus source and visualise it with the following graph (it is possible to do within one pipe).

```{r}
```

### 1.2
Calculate the ratio of b-words ($\frac{b\text{-words}}{b\text{-words}+r\text{-words}}$) for each stimulus and visualise ratio distribution using `geom_histogram`:
```{r}
```

### 1.3
Calculate the ratio of b-words ($\frac{b\text{-words}}{b\text{-words}+r\text{-words}}$) for each stimulus. Which stimulae have value between 0.5 and 0.6?
```{r, eval = FALSE}
```

### 2.

This dataset is based on (Coretta 2017, https://goo.gl/NrfgJm). This dissertation dealt with the relation between vowel duration and aspiration in consonants. Author carried out a data collection with 5 natives speakers of Icelandic. Then he extracted the duration of vowels followed by aspirated versus non-aspirated consonants. Check out whether vowels before aspirated consonants (like in Icelandic takka ‘key’ [tʰaʰka]) are signiﬁcantly shorter than vowels followed by non-aspirated consonants (like in kagga ‘barrel’ [kʰakka]).

Data are [here](http://math-info.hse.ru/f/2018-19/ling-data/icelandic.csv).

### 2.1 
Calculate a 95% confidence interval and a mean vowel duration (variable `vowel.dur`) for different contexts (variable `aspiration`) for all speakers (variable `speaker`) and visualise them using `geom_pointrange`.

```{r}
```



