---
title: "Confidence intervals in R"
output:
  html_document:
    df_print: paged
---

### Confidence intervals in R

Install library `DescTools` and load it:

```{r, eval=FALSE}
install.packages("DescTools")
```

```{r, warning=FALSE, message=FALSE}
library(DescTools)
```

### Confidence intervals for proportions

First, let us consider an abstract example so as to look at different effects connected with confidence intervals (sample size effect and confidence level effect). Suppose we tossed a coin 20 times and got 4 heads. 
```{r}
nheads <- 4 # number of heads
n1 <- 20  # total number of tosses
```

Now let's calculate a 95% confidence interval for the proportion of heads in such an experiment.

```{r}
BinomCI(nheads, n1) # 95% by default
```

Calculate the length of a confidence interval: 

```{r}
ci.95 <- BinomCI(nheads, n1)
ci.95[3] - ci.95[2]
```

Increase the number of tosses (number of heads remain the same):

```{r}
n2 <- 40 # now 40 tosses
ci.95.2 <- BinomCI(nheads, n2)
ci.95.2[3] - ci.95.2[2]  # it shrinked
```

Keep the number of tosses equal to 20, but increase the confidence level:

```{r}
ci.99 <- BinomCI(nheads, n1, conf.level = 0.99)
ci.99[3] - ci.99[2] # it extended
```

Now let's proceed to real data and work with *Verses* data set. 

```{r}
verses <- read.csv("https://raw.githubusercontent.com/LingData2019/LingData/master/data/poetry_last_in_lines.csv", sep = "\t")
str(verses) # recall which variables are there
```

Calculate a confidence interval for the proportion of nouns at the end of lines:

```{r}
nnouns <- nrow(verses[verses$UPoS == "NOUN", ])
total <- nrow(verses)

BinomCI(nnouns, total)
```

### Confidence intervals for means

Now let's work with a data set on Icelandic language from our previuos class.

```{r}
phono <- read.csv("http://math-info.hse.ru/f/2018-19/ling-data/icelandic.csv")
```

Choose aspirated and non-aspirated cases again:

```{r}
asp <- phono[phono$aspiration == "yes", ]
nasp <- phono[phono$aspiration == "no", ]
```

Calculate confidence intervals for mean values of vowel duration in each group:

```{r}
MeanCI(asp$vowel.dur)
MeanCI(nasp$vowel.dur)
```

Plot them using `sciplot`:

```{r, warning=FALSE, message=FALSE}
# install.packages("sciplot")
library(sciplot)
lineplot.CI(data = phono, 
            response = vowel.dur, 
            x.factor = aspiration)
```


### Confidence intervals and significance of differences

* If two CI's for a population parameter (proportion, mean, median, etc) do not overlap, it means that true values of population parameters are significantly different.

* If two CI's for a population parameter overlap, true values of population parameters are likely to coincide (to be equal to each other), but **not** necessarily do so. For example, if two confidence intervals for means overlap, we cannot make a definite conclusion, more accurate testing is required (t-test). So, in general, comparison of confidence intervals (with the same confidence level, of course) is **not** equivalent to hypotheses testing.

Consider a case when two CI's for means overlap, but population means are significantly different. Let's select only cases with aspirated consonants and compare the average vowel duration for round and unrounded vowels.

```{r}
w1 <- phono[phono$aspiration == 'yes' & phono$roundness == "round", ]
w2 <- phono[phono$aspiration == 'yes' & phono$roundness == "unrounded", ]
```

Do CI's overlap?

```{r}
MeanCI(w1$vowel.dur)
MeanCI(w2$vowel.dur)
```

Can we conclude that mean vowel duration is different for round and unrounded vowels? 

Now: perform an accurate test, a two sample Student's t-test.

```{r}
# reject or not reject H0
t.test(w1$vowel.dur, w2$vowel.dur)
```

Null hypotheses should be rejected, so population means are different. 

Actually, testing hypothesis about the equality of population means is equivalent to finding whether *a CI for the difference of means* includes zero. 

```{r}
# CI for difference between means
MeanDiffCI(w1$vowel.dur, w2$vowel.dur)
```

So, intersection of CI's for means (or for any population parameters) $\ne$ CI for the difference includes zero $\ne$ $H_0$ about equality should not be rejected.

